### 2015-01-09 prelim/oral_prelim.tex
revisions and plots for prelim

\begin{frame}{Motivation}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Average behavior is important to understand, but it does not paint the whole picture
    \begin{itemize}
      \item e.g. When constructing river levees, engineers need to be able to estimate a 100-year or 1000-year flood levels
      \item e.g. Probability of ambient air pollution exceeding a certain threshold level
    \end{itemize}
    \item Estimating the probability of rare events is challenging because these events are, by definition, rare
    \item Spatial extremes is promising because it borrows informaation across space
    \item Spatial extremes is also useful for estimating probability of extremes at sites without data
  \end{itemize}
\end{frame}

@@

\begin{frame}{Motivation}
\begin{adjustwidth}{1em}{0em}
  Ozone compliance for Clean Air Act (EPA) \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Annual fourth-highest daily maximum 8-hour concentration, averaged over 3 years, not to exceed 75 ppb
    \item Annual fourth-highest is the 99th percentile for the year
    \item Common objectives are \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item To interpolate to unmonitored sites
      \item Detect changes in extremes over time
      \item Study meterological conditions that lead to extreme events
    \end{itemize}
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Non-spatial analysis: Block maxima}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $X_1, \ldots, X_n$ be i.i.d.
    \item Consider the block maximum $M_n = \max(X_1, \ldots, X_n)$
    \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
    \begin{align*}
      \frac{ M_n - b_n }{ a_n } \converged G(z)
    \end{align*}
    then $G(z)$ follows a generalized extreme value distribution (GEV) (Falk et al., 2011)
    \item This motivates the use of the GEV for block maximum data
  \end{itemize}
\end{frame}

@@

\begin{frame}{Non-spatial analysis: Peaks over threshold}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $X_1, \ldots, X_n \sim F$
    \item If there exist normalizing sequences $a_t > 0$ and $b_t \in \calR$ such that for any $x \ge 0$, as $T \rightarrow \infty$
    \begin{align*}
      \Pr\left(\frac{X - b_t}{a_t} > x \mid X > T\right) \converged H(x),
    \end{align*}
    where $T$ is a thresholding value, then $H(x)$ follows a generalized Pareto distribution (GPD) (Balkema and de Haan, 1974)
  \end{itemize}
\end{frame}

@@

\begin{frame}{Max-stable processes for spatial data}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Consider i.i.d. spatial processes $x_j(\bs)$, $j = 1, \ldots, J$
    \item Let $M_J(\bs) = \bigvee_{j=1}^J x_j(\bs_i)$ be the block maximum at site $\bs$
    \item If there exists normalizing sequences $a_J(\bs)$ and $b_J(\bs)$ such that for all sites, $\bs_i, i = 1, \ldots, d$,
    \begin{align*}
      a_J^{-1}(\bs) \left\{ M_J(\bs) - b_J(\bs) \right\} \converged Y(\bs)
    \end{align*}
    which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process
    \item Therefore, max-stable processes are the standard model for block maxima
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate representations}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Marginally at each site, observations follow a GEV distribution
    \item For a finite collection of sites the multivariate representation for the GEV (mGEV) is
    \begin{align*}
      \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
            V(\bz)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
      \item $H$ is a probability measure on $\Delta_d$
      \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate GEV challenges}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Only a few closed-form expressions for $V(\bz)$ exist (Stephenson, 2003)
    \item Two common forms for $V(\bz)$:
    \begin{itemize}
      \item Symmetric logistic
      \begin{align*}
        V(\bz) = \left[\sum_{i = 1}^n \left( \frac{ 1 }{ z_i } \right)^{1/\alpha}\right]^\alpha
      \end{align*}
      \item Asymmetric logistic
      \begin{align*}
        V(\bz) = \sum_{l = 1}^L \left[\sum_{i = 1}^n \left(\frac{w_{il}}{z_i} \right)^{1 / \alpha_l} \right]^{\alpha_l}
      \end{align*}
      where $w_{il} \in [0, 1]$ and $\sum_{l = 1}^L w_{il} = 1$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Extremal dependence: $\chi$ statistic}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Correlation is the most common measure of dependence
    \begin{itemize}
      \item Focuses on the center and not tails
      \item This makes it irrelevant for extreme value analysis
    \end{itemize}
    \item Extreme value analysis focuses on the $\chi$ statistic, a measure of extremal dependence given by
    \begin{align*}
      \chi(h) = \lim_{c \rightarrow \infty}\Pr[Y(\bs) > c \mid Y(\bt) > c]
    \end{align*}
    where $h = ||\bs - \bt||$
    \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $h$
  \end{itemize}
\end{frame}

@@

\begin{frame}{Three principal contributions}
\begin{adjustwidth}{1em}{0em}
  \begin{enumerate}[1.] \setlength{\itemsep}{1.5em}
    \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
    \item Predicting rare binary events with a spatially dependent generalized extreme-value link function
    \item A Bayesian hierarchical model to allow for non-stationary covariance in extreme value models
  \end{enumerate}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Gaussian spatial model}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item In geostatistics, $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
    \item Model properties
    \begin{itemize}
      \item Nice computing properties (closed-form likelihood)
      \item For a Gaussian spatial model $\chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
      \item Tail is not flexible
      \begin{itemize}
        \item Light-tailed
        \item Symmetric
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
\begin{adjustwidth}{1em}{0em}
  A more flexible alternative is the spatial skew-$t$ process (Zhang and El-Shaarawi, 2012)
  \begin{align*}
    Y(\bs) = \bX(\bs)\bbeta + \lambda |z| + v(\bs)
  \end{align*}
  where
  \begin{itemize} \setlength{\itemsep}{0.25em}
    \item $\lambda \in \calR$ controls the skewness
    \item $z \sim N(0, \sigma^2)$ is a random effect
    \item $v(\bs)$ is a Gaussian process with variance $\sigma^2$ and \Matern correlation
    \item $\sigma^2 \sim \text{IG}(a, b)$
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution}
\begin{adjustwidth}{1em}{0em}
  \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
  \begin{align*}
    Y(\bs) \sim \text{skew-t}(\bX(\bs), \bOmega, \alpha, \text{df}=2a)
  \end{align*}
  where
  \begin{itemize} \setlength{\itemsep}{0.25em}
    \item $\bX(\bs) \bbeta$ is the location
    \item $\bOmega = \frac{1}{ab} \bar{\bOmega}$ is a correlation matrix
    \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
    \item $\bSigma$ is a postive definite correlation matrix
    \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution properties}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Model properties: \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Flexible tail
      \begin{itemize}
        \item Skewness controlled by $\lambda$
        \item Weight of tails controlled by degrees of freedom $2a$
      \end{itemize}
      \item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011)
      \item Computation that is on the order of Gaussian computation
    \end{itemize}
    \item Challenge: Long-range dependence occurs because all observations (near and far) share the same $z$ and $\sigma^2$ (Padoan, 2011)
    \begin{align*}
      \lim_{h \rightarrow \infty} \chi(h) > 0
    \end{align*}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Proof that $\lim_{h \rightarrow \infty} \chi(h) = 0$}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $A$ be the area between two sites $\bs$ and $\bt$
    \item Let $N(A)$ be the number of knots in $A$
    \item Assume that $N(A) \sim \text{HPP}[\mu(A)]$, where \vspace{0.25em}
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item HPP$[\mu]$ is a homogeneous poisson process with intensity $\mu$
      \item $\mu(\cdot)$ is an intensity measure defined on $A$
      \item $\lim_{h \rightarrow \infty} \mu(A) = \infty$
    \end{itemize}
    \begin{align*}
      \Pr[N(A) = k] = \frac{ \mu(A)^k \exp\{-\mu(A)\}}{k!}
    \end{align*}
  \end{itemize}
\end{frame}

@@

\begin{frame}{MCMC details}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Three main steps \vspace{0.5em}
    \begin{enumerate}[1.] \setlength{\itemsep}{0.5em}
      \item Impute censored data below $T$
      \item Update parameters with standard random walk Metropolis Hastings or Gibbs sampling
      \item Make spatial predictions
    \end{enumerate}
    \item Priors are selected to be conjugate when possible
  \end{itemize}
\end{frame}

@@

\begin{frame}{Simulation study}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item 50 datasets for each setting \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item 144 sites in $[0, 10] \times [0, 10]$ \vspace{0.25em}
      \begin{itemize} \setlength{\itemsep}{0.25em}
        \item 100 training
        \item 44 testing
      \end{itemize}
    \end{itemize}
    \item Model parameters \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Spatial range: $\rho = 1$
      \item Skew parameter: $\lambda = 3$
      \item Degrees of freedom: 6 for $t$ distributions
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Brier score}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Brier score (Gneiting and Raftery, 2007) used to compare fits
    \item The Brier score for predicting exceedance of threshold $c$ is
    \begin{align*}
      [e(c) - P(c)]^2
    \end{align*}
    where \vspace{0.25em}
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $y$ is a test set value
      \item $e(c) = I[y > c]$
      \item $P(c)$ is the predicted probability of exceeding $c$
    \end{itemize}
    \item Relative Brier scores:
    \begin{align*}
      \text{BS}_\text{rel} = \frac{ \text{BS}_\text{method}}{ \text{BS}_\text{Gaussian}}
    \end{align*}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item 9 different analysis methods incorporating \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
      \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
      \item 4 threshold levels \vspace{0.25em}
      \begin{itemize} \setlength{\itemsep}{0.25em}
         \item $T = 0$
         \item $T = 50$ppb, $q(0.48)$
         \item $T = 75$ppb, $q(0.92)$
         \item $T = 85$ppb, $q(0.97)$
      \end{itemize}
    \end{itemize}
    \item Compare Brier scores using two-fold cross validation
  \end{itemize}
\end{frame}

@@

@@

\begin{frame}{Rare binary regression}
\begin{adjustwidth}{1em}{0em}
  Common examples with non-spatial analysis \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Logistic regression
    \begin{align*}
      \Pr[Y_i = 1] = \frac{ \exp(\bX_i \bbeta) }{ 1 + \exp(\bX_i \bbeta)}
    \end{align*}
    \item Probit regression
    \begin{align*}
      \Pr[Y_i = 1] = \Phi(\bX_i \bbeta)
    \end{align*}
    where $\Phi$ is the standard normal distribution function
    \item Cloglog regression
    \begin{align*}
      \Pr[Y_i = 1] = 1 - \exp[ -\exp(\bX_i \bbeta)]
    \end{align*}
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Rare spatial binary regression}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item We model $Y_i = I(Z_i > 0)$ where $Z_i\sim$ multivariate GEV (mGEV)  is a latent variable
    \item Hierarchical model for mGEV (Reich and Shaby, 2012)
    \begin{align*}
      Z(\bs) = U(\bs) \theta(\bs)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $U(\bs) \iid$ GEV(1, $\alpha$, $\alpha$) is a nugget effect
      \item $\theta(\bs) = [\sum_{l = 1}^L A_l w_l(\bs)^{1/\alpha}]^\alpha$ is the spatial process
      \item $\alpha \in (0, 1)$ controls strength of nugget relative to spatial dependence
      \item $A_l \iid$ Positive Stable($\alpha$) is a random effect representing the intensity
      \item $w_l(\bs)$ gives the weight of the intensity of the $l$th random effect on site $\bs$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

% TODO: more about this model for sure
\begin{frame}{Non-stationary covariance for extreme values}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Stationary covariance functions are a function of distance between two sites.
    \begin{align*}
      \rho[Y(\bs), Y(\bt)] = \rho(h)
    \end{align*}
    where $h = ||\bs - \bt||$
    \item This assumes the covariance is the same everywhere, e.g. east vs west, mountains vs desert
    \item Misspecifying the covariance can impair spatial prediction and statistical inference
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{Motivation}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Average behavior is important to understand, but it does not paint the whole picture
    \begin{itemize}
      \item e.g. When constructing river levees, engineers need to be able to estimate a 100-year or 1000-year flood levels
      \item e.g. Probability of ambient air pollution exceeding a certain threshold level
    \end{itemize}
    \item Estimating the probability of rare events is challenging because these events are, by definition, rare
    \item Spatial extremes is promising because it borrows information across space
    \item Spatial extremes is also useful for estimating probability of extremes at sites without data
  \end{itemize}
\end{frame}

@@

\begin{frame}{Motivation}
\begin{adjustwidth}{1em}{0em}
  Ozone compliance for Clean Air Act (EPA) \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Annual fourth-highest daily maximum 8-hour concentration, averaged over 3 years, not to exceed 75 ppb
    \item Annual fourth-highest is the 99th percentile for the year
    \item Common objectives are \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item To interpolate to unmonitored sites
      \item Detect changes in extremes over time
      \item Study meteorological conditions that lead to extreme events
    \end{itemize}
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Non-spatial analysis: Block maxima}
\begin{adjustwidth}{1em}{0em}
  Fisher-Tippett-Gnedenko \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $X_1, \ldots, X_n$ be i.i.d.
    \item Consider the block maximum $M_n = \max(X_1, \ldots, X_n)$
    \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
    \begin{align*}
      \frac{ M_n - b_n }{ a_n } \converged G(z)
    \end{align*}
    then $G(z)$ follows a generalized extreme value distribution (GEV) (Gnedenko, 1943)
    \item This motivates the use of the GEV for block maximum data
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Non-spatial analysis: Peaks over threshold}
\begin{adjustwidth}{1em}{0em}
  Pickands-Balkema-de Haan theorem \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $X_1, \ldots, X_n \iid F$
    \item If there exist normalizing sequences $a_T > 0$ and $b_T \in \calR$ such that for any $x \ge 0$, as $T \rightarrow \infty$
    \begin{align*}
      \Pr\left(\frac{X - b_T}{a_T} > x \mid X > T\right) \converged H(x),
    \end{align*}
    where $T$ is a thresholding value, then $H(x)$ follows a generalized Pareto distribution (GPD) (Balkema and de Haan, 1974)
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Max-stable processes for spatial data}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Consider i.i.d. spatial processes $x_j(\bs)$, $j = 1, \ldots, J$
    \item Let $M_J(\bs) = \bigvee_{j=1}^J x_j(\bs_i)$ be the block maximum at site $\bs$
    \item If there exists normalizing sequences $a_J(\bs)$ and $b_J(\bs)$ such that for all sites, $\bs_i, i = 1, \ldots, d$,
    \begin{align*}
      \frac{M_J(\bs) - b_J(\bs)}{a_J(\bs)} \converged G(\bs)
    \end{align*}
    then $G(\bs)$ is a max-stable process (Smith, 1990)
    \item Therefore, max-stable processes are the standard model for block maxima
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate representations}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Marginally at each site, observations follow a GEV distribution
    \item For a finite collection of sites the multivariate representation for the GEV (mGEV) is
    \begin{align*}
      \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp[-V(\bz)]\\
            V(\bz)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $V(\bz)$ is called the exponent measure
      \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
      \item $H$ is a probability measure on $\Delta_d$
      \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate GEV challenges}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Only a few closed-form expressions for $V(\bz)$ exist
    \item Two common forms for $V(\bz)$
    \begin{itemize}
      \item Symmetric logistic (Gumbel, 1960)
      \begin{align*}
        V(\bz) = \left[\sum_{i = 1}^n \left( \frac{ 1 }{ z_i } \right)^{1/\alpha}\right]^\alpha
      \end{align*}
      \item Asymmetric logistic (Coles and Tawn, 1991)
      \begin{align*}
        V(\bz) = \sum_{l = 1}^L \left[\sum_{i = 1}^n \left(\frac{w_{il}}{z_i} \right)^{1 / \alpha_l} \right]^{\alpha_l}
      \end{align*}
      where $w_{il} \in [0, 1]$ and $\sum_{l = 1}^L w_{il} = 1$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Extremal dependence: $\chi$ statistic}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Correlation is the most common measure of dependence
    \begin{itemize}
      \item Focuses on the center and not tails
      \item This makes it irrelevant for extreme value analysis
    \end{itemize}
    \item Extreme value analysis focuses on the $\chi$ statistic (Coles et al., 1999), a measure of extremal dependence given by
    \begin{align*}
      \chi(h) = \lim_{c \rightarrow \infty}\Pr[Y(\bs) > c \mid Y(\bt) > c]
    \end{align*}
    where $h = ||\bs - \bt||$
    \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $h$
  \end{itemize}
\end{frame}

@@

\begin{frame}{Three principal contributions}
\begin{adjustwidth}{1em}{0em}
  \begin{enumerate}[1.] \setlength{\itemsep}{1.5em}
    \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
    \item Predicting rare binary events with a spatially dependent generalized extreme value link function
    \item A Bayesian hierarchical model to allow for non-stationary dependence in extreme value models
  \end{enumerate}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Gaussian spatial model}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item In geostatistics, $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$
    \item Model properties
    \begin{itemize}
      \item Nice computing properties (closed-form likelihood)
      \item For a Gaussian spatial model $\chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
      \item Tail is not flexible
      \begin{itemize}
        \item Light-tailed
        \item Symmetric
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
\begin{adjustwidth}{1em}{0em}
  A more flexible alternative is the spatial skew-$t$ process (Zhang and El-Shaarawi, 2012)
  \begin{align*}
    Y(\bs) = \bX(\bs)\bbeta + \lambda |z| + v(\bs)
  \end{align*}
  where
  \begin{itemize} \setlength{\itemsep}{0.25em}
    \item $\lambda \in \calR$ controls the skewness
    \item $z \sim N(0, \sigma^2)$ is a random effect
    \item $v(\bs)$ is a Gaussian process with variance $\sigma^2$, \Matern correlation, and $\gamma$ is the proportion of spatial variation in the \Matern correlation
    \item $\sigma^2 \sim \text{IG}(a, b)$
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution}
\begin{adjustwidth}{1em}{0em}
  \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
  \begin{align*}
    Y(\bs) \sim \text{skew-t}(\bX(\bs) \bbeta, \bOmega, \alpha, \text{df}=2a)
  \end{align*}
  where
  \begin{itemize} \setlength{\itemsep}{0.25em}
    \item $\bX(\bs) \bbeta$ is the location
    \item $\bOmega = \frac{1}{ab} \bar{\bOmega}$ is a correlation matrix
    \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
    \item $\bSigma$ is a positive definite correlation matrix
    \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution properties}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Model properties: \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Flexible tail
      \begin{itemize}
        \item Skewness controlled by $\lambda$
        \item Weight of tails controlled by degrees of freedom $2a$
      \end{itemize}
      \item Computation that is on the order of Gaussian computation
    \end{itemize}
    \item Challenge: For a skew-$t$ distribution $\lim_{h \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011) \vspace{-1em}
    \begin{itemize}
      \item Long-range dependence occurs because all observations (near and far) share the same $z$ and $\sigma^2$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Proof that $\lim_{h \rightarrow \infty} \chi(h) = 0$}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $A$ be the area between two sites $\bs$ and $\bt$
    \item Let $N(A)$ be the number of knots in $A$
    \item Assume that $N(A) \sim \text{HPP}[\mu(A)]$, where \vspace{0.25em}
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item HPP$[\mu(A)]$ is a homogeneous Poisson process with intensity measure $\mu(A)$ defined on $A$
      \item $\lim_{h \rightarrow \infty} \mu(A) = \infty$
    \end{itemize}
    \begin{align*}
      \Pr[N(A) = k] = \frac{ \mu(A)^k \exp\{-\mu(A)\}}{k!}
    \end{align*}
  \end{itemize}
\end{frame}

@@

\begin{frame}{MCMC details}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Three main steps \vspace{0.5em}
    \begin{enumerate}[1.] \setlength{\itemsep}{0.5em}
      \item Impute censored data below $T$
      \item Update parameters with Metropolis Hastings or Gibbs sampling
      \item Make spatial predictions
    \end{enumerate}
    \item Priors are selected to be conjugate when possible
  \end{itemize}
\end{frame}

@@

\begin{frame}{Simulation study}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item 50 datasets for each setting \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item 144 sites in $[0, 10] \times [0, 10]$ \vspace{0.25em}
      \begin{itemize} \setlength{\itemsep}{0.25em}
        \item 100 training
        \item 44 testing
      \end{itemize}
    \end{itemize}
    \item Model parameters \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Spatial range: $\rho = 1$
      \item Skew parameter: $\lambda = 3$
      \item Degrees of freedom: 6 for $t$ distributions
      \item Proportion of spatial variation: $\gamma = 0.9$
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Brier score}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Brier score (Gneiting and Raftery, 2007) used to compare fits
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Lower scores indicate better fits
    \end{itemize}
    \item The Brier score for predicting exceedance of threshold $c$ is
    \begin{align*}
      [e(c) - P(c)]^2
    \end{align*}
    where \vspace{0.25em}
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $y$ is a test set value
      \item $e(c) = I[y > c]$
      \item $P(c)$ is the predicted probability of exceeding $c$
    \end{itemize}
    \item Relative Brier scores:
    \begin{align*}
      \text{BS}_\text{rel} = \frac{ \text{BS}_\text{method}}{ \text{BS}_\text{Gaussian}}
    \end{align*}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Many different analysis methods incorporating \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Gaussian, symmetric-$t$, skew-$t$, and max-stable marginal distributions
      \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
      \item 4 threshold levels for $t$ marginals \vspace{0.25em}
      \begin{itemize} \setlength{\itemsep}{0.25em}
         \item $T = 0$
         \item $T = 50$ppb, $q(0.48)$
         \item $T = 75$ppb, $q(0.92)$
         \item $T = 85$ppb, $q(0.97)$
      \end{itemize}
      \item Thresholded at $T = 75$ for max-stable
    \end{itemize}
    \item Compare Brier scores from two-fold cross validation
  \end{itemize}
\end{frame}

@@

\begin{frame}{Probability of exceedance}
  \centering
  \begin{figure}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-gaus.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-t1.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-t6.pdf}
    \caption{Probability of exceeding 75ppb at least two days for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
  \end{figure}
\end{frame}

\begin{frame}{Quantile plots}
  \centering
  \begin{figure}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-gaus.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-t1.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-t6.pdf}
    \caption{$\widehat{q}(0.95)$ for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
  \end{figure}
\end{frame}

\begin{frame}{Quantile plots}
  \centering
  \begin{figure}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-gaus.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-t1.pdf}
    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-t6.pdf}
    \caption{$\widehat{q}(0.99)$ for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
  \end{figure}
\end{frame}

@@

\begin{frame}{Rare binary regression}
\begin{adjustwidth}{1em}{0em}
  Common examples with non-spatial analysis \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Logistic regression
    \begin{align*}
      \Pr[Y_i = 1] = \frac{ \exp(\bX_i \bbeta) }{ 1 + \exp(\bX_i \bbeta)}
    \end{align*}
    \item Probit regression
    \begin{align*}
      \Pr[Y_i = 1] = \Phi(\bX_i \bbeta)
    \end{align*}
    where $\Phi$ is the standard normal distribution function
    % \item Cloglog regression
    % \begin{align*}
    %   \Pr[Y_i = 1] = 1 - \exp[ -\exp(\bX_i \bbeta)]
    % \end{align*}
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Rare spatial binary regression}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item We model $Y_i = I(Z_i > 0)$ where $Z_i\sim$ multivariate GEV (mGEV)  is a latent variable
    \item Hierarchical model for mGEV (Reich and Shaby, 2012)
    \begin{align*}
      Z(\bs) = U(\bs) \theta(\bs)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $U(\bs) \iid$ GEV(1, $\alpha$, $\alpha$) is a nugget effect
      \item $\theta(\bs) = [\sum_{l = 1}^L A_l w_l(\bs)^{1/\alpha}]^\alpha$ is the spatial process
      \item $A_l \iid$ Positive Stable($\alpha$) is a random effect representing the intensity
      \item $w_l(\bs)$ gives the weight of the intensity of the $l$th random effect on site $\bs$
      \item $\alpha \in (0, 1)$ controls strength of nugget relative to spatial dependence
    \end{itemize}
  \end{itemize}
\end{frame}

@@

% TODO: more about this model for sure
\begin{frame}{Non-stationary dependence for extreme values}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Stationary covariance functions are a function of distance between two sites.
    \begin{align*}
      \rho[Y(\bs), Y(\bt)] = \rho(h)
    \end{align*}
    where $h = ||\bs - \bt||$
    \item This assumes the covariance is the same everywhere, e.g. east vs west, mountains vs desert
    \item Misspecifying the covariance can impair spatial prediction and statistical inference
  \end{itemize}
\end{frame}

\begin{frame}{Non-stationary dependence for extreme values}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item In extremes, stationary extremal dependence means
    \begin{align*}
      \chi(h) = \Pr[Y(\bs) > c | Y(\bt) > c]
    \end{align*}
    \item Currently, there are no methods to model non-stationarity in spatial extremes
    \item Semiparametric approach using spectral density ratios (de Carvalho and Davison, 2014)
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item Vector of observations can be transformed to pseudo-polar coordinates
      \item Pairwise analysis
    \end{itemize}
    \item New approach extending Reich and Shaby (2012)
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item Current model uses a single bandwidth term $\rho$ for all knots
      \item Proposed idea is to implement a knot-specific $\rho$ to induce non-stationarity
    \end{itemize}
  \end{itemize}
\end{frame}
